Appendix

### 2021-08-26 paper/appendix.Rmd
Clean appendix

```{r}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
```

```{r, echo = FALSE, message = FALSE, collapse = FALSE}
library(drake)
library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)

@@

\beginsupplement

@@

## Maps

@@

## Range of variation

@@

```{r}
cor_cap <- paste0("Correlation plot among variables for spatial and temporal
  relationships. ", "Median values over temporal series in each station were
  used for spatial relationships. ", "For temporal relationships, we used the
  temporal trends in each station. Futhermore, each station was weighted by the
  average error standard on the temporal trends of the variables.")
```

@@

set_names_corr <- function(mat = NULL) {
  colnames(mat) <- str_replace_all(colnames(mat), var_replacement())
  rownames(mat) <- str_replace_all(rownames(mat), var_replacement())
  return(mat)
}

corr_plot_tps %<>% set_names_corr(.)
corr_plot_med %<>% set_names_corr(.)

par(mfrow = c(1, 2))
corrplot::corrplot(corr_plot_med, type = "upper", diag = FALSE, order =
  "alphabet")
title("Spatial")
corrplot::corrplot(corr_plot_tps, type = "upper", diag = FALSE, order =
  "alphabet")
title("Temporal trends")
```

@@

## Temporal model with richness and biomass

@@

We modeled the temporal trends of food-web structure as dependant of the
temporal trends of both species richness and community biomass as:

$dN = dB + dR + \epsilon$

@@

```{r sens, out.width="100%"}
# Merge reg_table
merged_reg_table <- pmap_dfr(
  list(
    df = list(
      reg_table_bm_rich_mono_trends,
      reg_table_bm_rich_mono_stable_trends,
      reg_table_bm_rich_trends,
      reg_table_bm_rich
      ),
    name = list(
      "Only monotonuous trends",
      "Monotonuous + stable trends",
      "All significant trends",
      "All trends"
      ),
    st = list(
      st_mono_trends_rich_bm,
      st_mono_trends_stable_rich_bm,
      st_trends_rich_bm,
      seq(1,400)
    )
    ), function (df, name, st) {
    df$selection <- paste0(name, " (N = ", length(st),")")
    return(select(df, selection, everything()))
  }
)

merged_reg_table %>%
  filter(! term %in% "(Intercept)", ! response %in% "piel_nind") %>%
  mutate(
   response = str_replace_all(response, var_replacement()),
   term = str_replace_all(term, var_replacement()),
  ) %>%
  ggplot(aes(x = term, y = std_estimate)) +
  geom_point(aes(color = selection)) +
  facet_grid(cols = vars(response)) +
  labs(
    x = "Explicative variable",
    y = "Standardized coefficients"
  ) +
  theme(legend.position = "bottom", axis.text.x = element_text(hjust=1, angle = 15))
```

@@

### ANOVA

```{r}
model_bm_rich_int_mono_stable_trends %>%
  .[names(.) %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")] %>%
  map(., car::Anova) %>%
  map(., tidy) %>%
  map2_dfr(., names(.), ~.x %>% mutate(response = .y) %>% select(response,
      everything())) %>%
  kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "ANOVA results with interactions."
      ),
    label = "aov-int"
    ) %>%
kable_styling(latex_options =c("scale_down", "stripped")) %>%
collapse_rows(columns = 1, valign = "top")
#map(model_bm_rich_int_mono_stable_trends, ~try(car::vif(.x)))
```

```{r}
model_bm_rich_int_mono_stable_trends %>%
  .[names(.) %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")] %>%
  map(., tidy) %>%
  map2_dfr(., names(.), ~.x %>% mutate(response = .y) %>% select(response,
      everything())) %>%
  kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "Regression results with interactions."
      ),
    label = "aov-int"
    ) %>%
kable_styling(latex_options =c("scale_down", "stripped")) %>%
collapse_rows(columns = 1, valign = "top")
```

### Plots

```{r}
pred_int_rich <- map(model_bm_rich_int_mono_stable_trends,
      ~ggpredict(.x,
        terms = c("log_rich_std", "log_bm_std [quart2]")) %>%
        as_tibble() %>%
        mutate(group = as.numeric(as.character(group))) %>%
        rename("log_rich_std" = x, "log_bm_std" = group)
        ) %>%
enframe(name = "response", value = "pred_int")

ti <- pred_bm_rich_mono_stable_trends %>%
  filter(term == "log_rich_std", response %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")) %>%
  select(response, raw_plot) %>%
  left_join(pred_int_rich) %>%
  mutate(pred_plot_color = pmap(
      list(raw_plot, response, pred_int),
      function(p, resp, pred) {
        p +
          geom_line(data = pred,
            aes(
              y = predicted,
              x = log_rich_std,
              color = as.factor(log_bm_std)
            ), size = 1
          ) +
        scale_color_discrete()
      }
      )
  )
rich_leg <- get_legend(ti$pred_plot_color[[1]])

plot_grid(plotlist = map(ti$pred_plot_color, ~.x + theme(legend.position = "none")), rich_leg)
```

@@

```{r}
pred_int_bm <- map(model_bm_rich_int_mono_stable_trends,
      ~ggpredict(.x,
        terms = c("log_bm_std", "log_rich_std [quart2]")) %>%
        as_tibble() %>%
        mutate(group = as.numeric(as.character(group))) %>%
        rename("log_bm_std" = x, "log_rich_std" = group)
        ) %>%
enframe(name = "response", value = "pred_int")

ti_bm <- pred_bm_rich_mono_stable_trends %>%
  filter(term == "log_bm_std", response %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")) %>%
  select(response, raw_plot) %>%
  left_join(pred_int_bm) %>%
  mutate(pred_plot_color = pmap(
      list(raw_plot, response, pred_int),
      function(p, resp, pred) {
        p +
          geom_line(data = pred,
            aes(
              y = predicted,
              x = log_bm_std,
              color = as.factor(log_rich_std)
            ), size = 1
          ) +
        scale_color_discrete()
      }
      )
  )

bm_leg <- get_legend(ti_bm$pred_plot_color[[1]])
plot_grid(plotlist = map(ti_bm$pred_plot_color, ~.x + theme(legend.position = "none")), bm_leg)
```

```{r, echo=FALSE, eval=FALSE}
test <- model_bm_rich_int_mono_stable_trends[["ct_ff"]]
car::Anova(test)
co <- coef(test)
pred <- ggpredict(
  test,
  terms = c("log_bm_std ", "log_rich_std")
  ) %>%
  as_tibble() %>%
  mutate(group = as.numeric(as.character(group))) %>%
  rename("log_bm_std" = x, "log_rich_std" = group)

@@

```{r, result=FALSE}
mod4_3d <- model_bm_rich_int_mono_stable_trends[
  names(model_bm_rich_int_mono_stable_trends) %in%
    c("ct_ff", "w_trph_lvl_avg")]

@@

```{r,result=FALSE}
mat <- reshape2::acast(
  pred_list$w_trph_lvl_avg,
  log_rich_std~log_bm_std,
  value.var = "y")
p <- plot_ly(
  x = ~as.numeric(colnames(mat)),
  y = ~as.numeric(rownames(mat)),
  z = ~mat, ) %>%
  add_surface() %>%
  layout(scene = list(
      xaxis = list(title = "Community biomass"),
      yaxis = list(title = "species richness"),
      zaxis = list(title = "Avg trophic level")
      ))
```

```{r,result=FALSE, eval=FALSE}
test <- model_bm_rich_int_mono_stable_trends[["piel_bm"]]
car::Anova(test)
co <- coef(test)
pred <- ggpredict(
  test,
  terms = c("log_bm_std ", "log_rich_std")
  ) %>%
  as_tibble() %>%
  mutate(group = as.numeric(as.character(group))) %>%
  rename("log_bm_std" = x, "log_rich_std" = group)

@@

```{r}
p_ct_ff <- ggdraw() +
    draw_image(get_mypath("figures", "int_rich_bm_cf_ff.png"), scale = 1)
p_tlvl <- ggdraw() +
    draw_image(get_mypath("figures", "int_rich_bm_tlvl.png"), scale = 1)
p_piel_bm <- ggdraw() +
    draw_image(get_mypath("figures", "int_rich_bm_eveness.png"), scale = 1)
plot_grid(
  p_ct_ff,
  p_tlvl,
  p_piel_bm
)
```

- Results for interactions: 
  - Connectance: 
    - Always decreases when richness decreases (inverse of expectation)
    - Increases when richness increases (inverse of expectation), but always
      decreases if biomass decreases
    - Increases when biomass increases
  - Average trophic level:
    - Always decreases when biomass decreases 
    - Always increases when biomass increases 
    - Whatever the richness dynamics, it is the biomass dynamics which drives
      all
  - Pielou of biomass:
    - Always positive when biomass decreases
    - Always negative when biomass increases

>>>>

```{r, echo=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
```

```{r, echo=FALSE, message=FALSE, results="hide"}
library(drake)
library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)

@@

\beginsupplement
\newpage

@@

@@

@@

```{r}
cor_cap <- paste0("Correlation plot among variables for spatial and temporal
  relationships. ", "Median values over temporal series in each station were
  used for spatial relationships. ", "For temporal relationships, we used the
  temporal trends in each station. Futhermore, each station was weighted by the
  average error standard on the temporal trends of the variables.",
  " Pisc: piscivorous, Nb: number, node: node of the food-web corresponding to a
  trophic species, prop: proportion."
)
```

@@

rv <- var_replacement()
rv_more <- c(
  nbnode_std = "Nb node",
  nb_pisc_rich_std = "Nb pisc species",
  nb_pisc_node_std = "Nb pisc nodes",
  prop_pisc_node = "Prop pisc node",
  prop_pisc_rich = "Prop pisc species"
)

rv <- c(rv_more, rv[!names(rv) == "rich_std"])

set_names_corr <- function(mat = NULL, replacement_vector = var_replacement()) {
  colnames(mat) <- str_replace_all(colnames(mat), replacement_vector)
  rownames(mat) <- str_replace_all(rownames(mat), replacement_vector)
  return(mat)
}

corr_plot_tps %<>% set_names_corr(., replacement_vector = rv)
corr_plot_med %<>% set_names_corr(., replacement_vector = rv)

#par(mfrow = c(1, 2))
corrplot::corrplot(
  corr_plot_med,
  type = "upper",
  diag = FALSE,
  order = "alphabet",
  tl.srt = 30,
  tl.cex = .7
)
title("Spatial", adj = 0, line = 3)
corrplot::corrplot(corr_plot_tps, type = "upper", diag = FALSE, order =
  "alphabet",
  tl.srt = 30,
  tl.cex = .7
)
title("Temporal trends", adj = 0, line = 3.2)
```

@@

## Sensisivity analysis for the temporal model

@@

```{r}
cap_sens_temp_model <- paste0("
  Testing the robustness of the estimated coefficients to the selection of
  sites. 'All trends' corresponds to all the sites, 'Only monotonuous trends'
  corresponds to all the trends negative or positive monotonically, 'Monotonuous
  + stable' includes the previous selection and the stable ones (the selection
    used in main text), 'All significant trends' correspond to all the trends
significantly negative or positive.", " We classified a site as
  having a significant trend when it displays both significant trends in biomass
and in species richness."
)
```
```{r sens, out.width="100%", fig.cap=cap_sens_temp_model}
# Merge reg_table
merged_reg_table <- pmap_dfr(
  list(
    df = list(
      reg_table_bm_rich_mono_trends,
      reg_table_bm_rich_mono_stable_trends,
      reg_table_bm_rich_trends,
      reg_table_bm_rich
      ),
    name = list(
      "Only monotonuous",
      "Monotonuous + stable (used in main text)",
      "All significant trends",
      "All sites"
      ),
    st = list(
      st_mono_trends_rich_bm,
      st_mono_trends_stable_rich_bm,
      st_trends_rich_bm,
      seq(1,400)
    )
    ), function (df, name, st) {
    df$selection <- paste0(name, " (N = ", length(st),")")
    return(select(df, selection, everything()))
  }
)

merged_reg_table %>%
  filter(! term %in% "(Intercept)", ! response %in% "piel_nind") %>%
  mutate(
   response = str_replace_all(response, var_replacement()),
   term = str_replace_all(term, var_replacement()),
  ) %>%
  ggplot(aes(x = term, y = std_estimate)) +
  geom_point(aes(color = selection)) +
  facet_grid(cols = vars(response)) +
  labs(
    x = "Explicative variable",
    y = "Standardized coefficients"
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(hjust = 1, angle = 15, size = 8),
    text = element_text(size = 8)) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))
```

We found that selection of data can have either low or high effects on estimated
effects. When selecting only the sites that display significant decrease or
increase of community biomass and species richness across time, i.e. 
a tenth of the site, we found stronger links among the temporal dynamics of
average trophic level, connectance, biomass and the temporal dynamics of biomass
and species richness (Fig. \@ref(fig:sens), purple and red points). On the other
side, we found that the data selection that we used in the main analysis
displays roughly the same results that the selection that includes all sites
(Fig. \@ref(fig:sens), blue and green points).

The fact that we observed differences with data selection is that  "Only
monotonuous" and "All significant trends" do not contain temporal dynamics
around 0 because they are often non significant. We think that this data
selection is not satisfactory because we wanted to include the temporal dynamics
of biomass of species richness that were not significant, i.e. that can be
considered as stable. The rational is that we wanted to know what was the
expected temporal trends of food-web structure while the ones of biomass and
species richness are kept stable. Without those data, the regression becomes
constrained only by the extreme temporal trends of species richness and
community biomass.

@@

@@

```{r}
model_bm_rich_int_mono_stable_trends %>%
  .[names(.) %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")] %>%
  map(., car::Anova) %>%
  map(., tidy) %>%
  map2_dfr(., names(.), ~.x %>% mutate(response = .y) %>% select(response,
      everything())) %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
  ) %>%
  rename_all(~str_to_title(.)) %>%
  kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "ANOVA results with interactions."
      ),
    label = "aov-int"
    ) %>%
kable_styling(latex_options =c("scale_down", "stripped")) %>%
collapse_rows(columns = 1, valign = "top")
#map(model_bm_rich_int_mono_stable_trends, ~try(car::vif(.x)))
```

```{r}
model_bm_rich_int_mono_stable_trends %>%
  .[names(.) %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")] %>%
  map(., tidy) %>%
  map2_dfr(., names(.), ~.x %>% mutate(response = .y) %>% select(response,
      everything())) %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
  ) %>%
  rename_all(~str_to_title(.)) %>%
  kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "Regression results of the model with interactions between the temporal
      dynamics of community biomass and the one of species richness."
      ),
    label = "reg-int"
    ) %>%
kable_styling(latex_options =c("scale_down", "stripped")) %>%
collapse_rows(columns = 1, valign = "top")
```

```{r, results="hide", echo=FALSE}
pred_int_rich <- map(model_bm_rich_int_mono_stable_trends,
      ~ggpredict(.x,
        terms = c("log_rich_std", "log_bm_std [quart2]")) %>%
        as_tibble() %>%
        mutate(group = as.numeric(as.character(group))) %>%
        rename("log_rich_std" = x, "log_bm_std" = group)
        ) %>%
enframe(name = "response", value = "pred_int")

ti <- pred_bm_rich_mono_stable_trends %>%
  filter(
    term == "log_rich_std",
    response %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")
    ) %>%
  select(response, raw_plot) %>%
  left_join(pred_int_rich) %>%
  mutate(pred_plot_color = pmap(
      list(raw_plot, response, pred_int),
      function(p, resp, pred, ...) {
        p +
          geom_line(data = pred,
            aes(
              y = predicted,
              x = log_rich_std,
              color = as.factor(log_bm_std)
            ), size = 1
          ) +
        scale_color_discrete() +
        labs(x = "Log species richness",
          y = var_replacement()[resp],
          color = "Log community biomass"
          ) +
        theme(
          text = element_text(size = 8),
          axis.text = element_text(size = 8)
        )
      }
        )
  )
rich_leg <- get_legend(ti$pred_plot_color[[1]])
```

```{r}
int_rich_cap <- paste0(
  "Relationships between the temporal dynamics of food-web
  structure and the one of species richness. The interaction between
  the temporal dynamic of species richness and community biomass is significant
  for the three descriptors of food-web structure. The chosen values for the
  community biomass temporal dynamic correspond respectively to the first,
  second and third quartiles.
  ")

```

```{r int-rich, fig.cap=int_rich_cap}
plot_grid(plotlist = 
  map(ti$pred_plot_color,
    ~.x + theme(
      legend.position = "none",
    )
    ),
rich_leg)
```

@@

```{r, results="hide", echo=FALSE}
pred_int_bm <- map(model_bm_rich_int_mono_stable_trends,
      ~ggpredict(.x,
        terms = c("log_bm_std", "log_rich_std [quart2]")) %>%
        as_tibble() %>%
        mutate(group = as.numeric(as.character(group))) %>%
        rename("log_bm_std" = x, "log_rich_std" = group)
        ) %>%
enframe(name = "response", value = "pred_int")

ti_bm <- pred_bm_rich_mono_stable_trends %>%
  filter(term == "log_bm_std", response %in% c("ct_ff", "w_trph_lvl_avg", "piel_bm")) %>%
  select(response, raw_plot) %>%
  left_join(pred_int_bm) %>%
  mutate(pred_plot_color = pmap(
      list(raw_plot, response, pred_int),
      function(p, resp, pred) {
        p +
          geom_line(data = pred,
            aes(
              y = predicted,
              x = log_bm_std,
              color = as.factor(log_rich_std)
            ), size = 1
          ) +
        scale_color_discrete() +
                labs(x = "Log species richness",
          y = var_replacement()[resp],
          color = "Log community biomass"
          ) +
        theme(
          text = element_text(size = 8),
          axis.text = element_text(size = 8)
        )

      }
      )
  )
bm_leg <- get_legend(ti_bm$pred_plot_color[[1]])
```

```{r}
int_bm_cap <- paste0(
  "Relationships between the temporal dynamics of food-web
  structure and the one of community biomass. The interaction between
  the temporal dynamics of species richness and community biomass is significant
  for the three descriptors of food-web structure. The chosen values for the
  species richness temporal dynamic correspond respectively to the first,
  second and third quartiles.
  "
)
```

```{r int-bm, fig.cap=int_bm_cap}
plot_grid(plotlist = map(ti_bm$pred_plot_color, ~.x + theme(legend.position = "none")), bm_leg)
```

```{r, echo=FALSE, results="hide"}
test <- model_bm_rich_int_mono_stable_trends[["ct_ff"]]
car::Anova(test)
co <- coef(test)
pred <- ggpredict(
  test,
  terms = c("log_bm_std ", "log_rich_std")
  ) %>%
  as_tibble() %>%
  mutate(group = as.numeric(as.character(group))) %>%
  rename("log_bm_std" = x, "log_rich_std" = group)

@@

```{r, results=FALSE}
mod4_3d <- model_bm_rich_int_mono_stable_trends[
  names(model_bm_rich_int_mono_stable_trends) %in%
    c("ct_ff", "w_trph_lvl_avg")]

@@

```{r,results=FALSE}
mat <- reshape2::acast(
  pred_list$w_trph_lvl_avg,
  log_rich_std~log_bm_std,
  value.var = "y")
p <- plot_ly(
  x = ~as.numeric(colnames(mat)),
  y = ~as.numeric(rownames(mat)),
  z = ~mat, ) %>%
  add_surface() %>%
  layout(scene = list(
      xaxis = list(title = "Community biomass"),
      yaxis = list(title = "species richness"),
      zaxis = list(title = "Avg trophic level")
      ))
```

```{r,results=FALSE, eval=FALSE}
test <- model_bm_rich_int_mono_stable_trends[["piel_bm"]]
car::Anova(test)
co <- coef(test)
pred <- ggpredict(
  test,
  terms = c("log_bm_std ", "log_rich_std")
  ) %>%
  as_tibble() %>%
  mutate(group = as.numeric(as.character(group))) %>%
  rename("log_bm_std" = x, "log_rich_std" = group)

@@

```{r}
int_3d_cap <- paste0(
  "Tri-dimentional plot displaying the relationships among food-web structure
  from one side and the temporal dynamics of species richness and community
  biomass from the other side. The surface represents the values predicted by
  the model."
)
```

```{r int-3d, fig.cap=int_3d_cap}
p_ct_ff <- ggdraw() +
    draw_image(get_mypath("figures", "int_rich_bm_cf_ff.png"), scale = 1)
p_tlvl <- ggdraw() +
    draw_image(get_mypath("figures", "int_rich_bm_tlvl.png"), scale = 1)
p_piel_bm <- ggdraw() +
    draw_image(get_mypath("figures", "int_rich_bm_eveness.png"), scale = 1)
plot_grid(
  p_ct_ff,
  p_tlvl,
  p_piel_bm,
  labels = "AUTO"
)
```

All together, we found that interactions between the temporal dynamics of
community biomass and species richness have no major effects on the bivariate
relationships (Fig. \@ref(fig:int-bm) and \@ref(fig:int-rich)), at least in
the range of values observed in the dataset (1st to 3rd quartiles).  It is more
complicated for connectance when exploring the full range of possible values
(Fig. \@ref(fig:int-3d) A). We found that connectance strongly decrease with
increasing of community biomass temporal trends when species richness temporal
trends is strongly negative, while we found the opposite when species richness temporal
trends is strongly positive. Furthermore we found that the temporal dynamic of
connectance had a strong positive relationship with the one of species richness
when community biomass increases through time while we found no relationship
when the temporal trend of community biomass was strongly negative.

- Results for interactions: 
  - Connectance: 
    - Always decreases when richness decreases (inverse of expectation)
    - Increases when richness increases (inverse of expectation), but always
      decreases if biomass decreases
    - Increases when biomass increases
  - Average trophic level:
    - Always decreases when biomass decreases 
    - Always increases when biomass increases 
    - Whatever the richness dynamics, it is the biomass dynamics which drives
      all
  - Pielou of biomass:
    - Always positive when biomass decreases
    - Always negative when biomass increases 
    - Whatever the richness dynamics, it is the biomass dynamics which drives
      all
